---
title: "CS-201 First Look at R"
author: "Amber Camp"
date: 8/26/26
format: html
editor: visual
toc: true
---

This is the first class session walking through code together!

We will be programming in R and in this class will primarily use `qmd` documents to keep things tidy and presentable. Each code chunk below is where you will run R code. Feel free to add notes as we go along.

## Navigating RStudio

By default, everyone should have the same four-pane layout:

1.  **Top left**: Source editor

2.  **Bottom left**: Console (and Terminal and Background Jobs)

3.  **Top Right**: Environment

4.  **Bottom Right**: Files, Plots, Help, etc.

### Customizing look & feel

Feel free to customize your RStudio.\
Go to Tools \> Global Options \> Appearance to select a color scheme you like.

I also personally like to enable these:

- Tools \> Global Options \> Code \> Insert matching parens/quotes

- Tools \> Global Options \> Console \> Show syntax highlighting in console input

I don't recommend rearranging your pane layout at this time.

## Running Code

Try running some code

### Load Library

```{r}
library(ggplot2)

```

## Introducing `ggplot2`

### Some Scripting

```{r}
#This is a comment 
quiz1 <- 88 #This is how to assign a value to a name
quiz2 <- 94
quiz3 <- 79

quiz1

#averaging
(quiz1+quiz2+quiz3)/3

#mean
mean(quiz1,quiz2,quiz3)#not correct just returns mean of first argument 
mean(quiz1+quiz2+quiz3)#not correct just adds 

mean(c(quiz1,quiz2,quiz3))#correct way to run mean

list <- c(quiz1,quiz2,quiz3)#assign a list to a variable

mean(list)#another way to get a correct mean

quiz3 <- 99

average

list <- c(quiz1,quiz2,quiz3)
average <- mean(list)
mean(c(quiz1,quiz2,quiz3))

```

### More scores 

```{r}
scores <- c(88,94,79,91,85)
scores


mean(scores)
max(scores)
min(scores)
length(scores)

summary(scores)

students <- ("a","b","c","d","e")
students

#bar plot
barplot(scores,name.arg = students)
```

```{}
```

## Extra Credit (1pt)

What result do you expect when you run this expression?

```{r}

```

What did you observe?

> Answer:

Why do you think this is happening? (You can look online for help)

> Answer:
