---
title: "CS-201 First Look at R"
author: "Amber Camp"
date: 8/26/26
format: html
editor: visual
toc: true
---

This is the first class session walking through code together!

We will be programming in R and in this class will primarily use `qmd` documents to keep things tidy and presentable. Each code chunk below is where you will run R code. Feel free to add notes as we go along.

## Navigating RStudio

By default, everyone should have the same four-pane layout:

1.  **Top left**: Source editor

2.  **Bottom left**: Console (and Terminal and Background Jobs)

3.  **Top Right**: Environment

4.  **Bottom Right**: Files, Plots, Help, etc.

### Customizing look & feel

Feel free to customize your RStudio.\
Go to Tools \> Global Options \> Appearance to select a color scheme you like.

I also personally like to enable these:

- Tools \> Global Options \> Code \> Insert matching parens/quotes

- Tools \> Global Options \> Console \> Show syntax highlighting in console input

I don't recommend rearranging your pane layout at this time.

## Running Code

Try running some code

### Load Library

```{r}
library(ggplot2)

```

## Introducing `ggplot2`

### Some Scripting

```{r}
#This is a comment 
quiz1 <- 88 #This is how to assign a value to a name
quiz2 <- 94
quiz3 <- 79

quiz1

#averaging
(quiz1+quiz2+quiz3)/3

#mean
mean(quiz1,quiz2,quiz3)#not correct just returns mean of first argument 
mean(quiz1+quiz2+quiz3)#not correct just adds 

mean(c(quiz1,quiz2,quiz3))#correct way to run mean

list <- c(quiz1,quiz2,quiz3)#assign a list to a variable

mean(list)#another way to get a correct mean

quiz3 <- 99

average

list <- c(quiz1,quiz2,quiz3)
average <- mean(list)
mean(c(quiz1,quiz2,quiz3))

```

### More scores

```{r}
scores <- c(88,94,79,91,85)
scores


mean(scores)
max(scores)
min(scores)
length(scores)

summary(scores)

students <- c("a", "b", "c", "d", "e")
students

#bar plot
barplot(scores,names.arg = students)
barplot(scores,names.arg = students, ylim = c(0, 100), main="TITLE", xlab= "pupils", y= "what they got")

```

```         
```

## Introducing `ggplot2`

#### Make a Data Frame

```{r}
grades_df <- data.frame(student = students, score = scores)

grades_df


summary(grades_df)

nrow(grades_df) #number of rows
ncol(grades_df) #number of clumns 
dim(grades_df) #number of dimensions
```

### Try GGplot

```{r}
install.packages("ggplot2")

ggplot(data = grades_df, aes(x = student, y = score))

ggplot(data = grades_df, aes(x = student, y = score, group = 1)) +
  geom_col(fill = "steelblue") +
  geom_point() +
  geom_line() +
  theme_minimal() +
  labs(title = "Best Graph", 
       subtitle = "Quiz Scores",
       x = "Students",
       y = "what they scored")


```

### Mutate the data frame

```{r}

library(tidyverse)

grades_df <- grades_df |> 
  mutate(study_hours = c(2, 5, 1, 4, 3))



```

### Make a scatter plot

```{r}

ggplot(data = grades_df, aes(x = study_hours, y = score)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "purple") +
  theme_dark() +
  labs(title = "Students study for higher grades",
       x = "study time",
       y = "grades") 
  


#cannot use fill(use color instead), dont use reorder




```

## Extra Credit (1pt)

What result do you expect when you run this expression?

```{r}

1 + 1 +1

quiz1 + quiz2

1 +1 + 1 - 3

0.1 + 0.1 + 0.1 - 0.2

options(scipen = 999)

print(0.1, 20)
print(0.3, 20)
print(0.5, 20)




```

What did you observe?

> Answer:

Why do you think this is happening? (You can look online for help)

> Answer:
